FLM for growth as function of SPEI

### 2021-02-17 analysis/paper/vital rate models.Rmd
changed diagnostic plots to quartile residuals for binomial models.

```{r}
s_post_cf <-
  bam(surv ~ 
        s(log_size_prev, bs = "cr") +
        # s(year) +
        # s(year_fac, bs = "re") +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_cf,
      method = "REML",
      select = TRUE,
      cluster = cl)
```

```{r}
# gam.check(s_post_cf) #not sure what to expect from any of these plots
appraise(s_post_cf)
# concurvity(s_post_cf, full = FALSE)
```

@@

```{r}
s_post_1ha <-
  bam(surv ~ 
        s(log_size_prev, bs = "cr") +
        # s(year_fac, bs = "re") +
        s(plot, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_1ha,
      method = "REML",
      select = TRUE,
      cluster = cl)
```

@@

@@

```{r}
g_post_cf <-
  bam(log_size ~ 
        s(log_size_prev, bs = "cr") + 
        # s(year_fac, bs = "re") + 
        s(plot, bs = "re") + #random effect of plot
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #unimodal-ish response to drought, but one knot per month lag.
          xt = list(bs = "cr")),
      # family = gaussian(link = "identity"),
      family = scat(link = "identity"), #like leptokurtic gaussian.
      data = demog2_post_cf,
      method = "REML",
      select = TRUE,
      cluster = cl)
```

@@

```{r}
appraise(f_post_cf)
# concurvity(f_post_cf, full = FALSE)
```

@@

```{r}
appraise(f_post_1ha)
```

>>>>

```{r}
s_post_cf <-
  bam(surv ~ 
        s(log_size_prev, bs = "cr") +
        s(plot, bs = "re") + #random intercept
        # s(plot, log_size_prev, bs = "re") + #random slope
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_cf,
      method = "REML",
      select = TRUE,
      cluster = cl)
```

```{r}
# appraise(s_post_cf)
#quantile residuals are better with discrete response variable.
library(statmod)
#histogram
plot(density(qresid(s_post_cf)))
#QQ plot
qqnorm(qresid(s_post_cf)); qqline(qresid(s_post_cf))
#Fitted vs. residuals
scatter.smooth(predict(s_post_cf, type = "response"), qresid(s_post_cf), col = "grey")
```

@@

```{r}
s_post_1ha <-
  bam(surv ~ 
        s(log_size_prev, bs = "cr") +
        s(plot, bs = "re") +
        # s(plot, log_size_prev, bs = "re") +
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #cubic response (at most) to drought, one knot per month lag.
          xt = list(bs = "cr")),
      family = binomial,
      data = demog_post_1ha,
      method = "REML",
      select = TRUE,
      cluster = cl)
```
```{r}
#histogram
plot(density(qresid(s_post_1ha)))
#QQ plot
qqnorm(qresid(s_post_1ha)); qqline(qresid(s_post_1ha))
#Fitted vs. residuals
scatter.smooth(predict(s_post_1ha, type = "response"), qresid(s_post_1ha), col = "grey")
```

@@

Emilio suggested plotting curves from all 1-ha plots just to visualize the variation. But the random effect of plot is not significant, so there is actually very little variation among plots in the model.

```{r}
new_data <-
  tidyr::expand(model.frame(s_post_1ha),
                nesting(plot),
                log_size_prev = seq(min(log_size_prev), max(log_size_prev), length.out = 50),
                spei_history = mean(spei_history), L = L)
m1pred <- bind_cols(new_data,
          as.data.frame(predict(s_post_1ha, newdata = new_data, se.fit = TRUE, type = "response")))
ggplot(m1pred, aes(x = log_size_prev, y = fit, group = plot)) + geom_line()
```

The wide confidence intervals are probably due to lower sample size at this size of plant.

@@

```{r}
g_post_cf <-
  bam(log_size ~ 
        s(log_size_prev, bs = "cr") + 
        s(plot, bs = "re") + #random effect of plot on intercept
        s(spei_history, L,
          bs = "cb",
          k = c(3, 35),  #unimodal-ish response to drought, but one knot per month lag.
          xt = list(bs = "cr")),
      # family = gaussian(link = "identity"),
      family = scat(link = "identity"), #like leptokurtic gaussian.
      data = demog2_post_cf,
      method = "REML",
      select = TRUE,
      cluster = cl)
```

@@

```{r}
#histogram
plot(density(qresid(f_post_cf)))
#QQ plot
qqnorm(qresid(f_post_cf)); qqline(qresid(f_post_cf))
#Fitted vs. residuals
scatter.smooth(predict(f_post_cf, type = "response"), qresid(f_post_cf), col = "grey")
```

@@

```{r}
#histogram
plot(density(qresid(f_post_1ha)))
#QQ plot
qqnorm(qresid(f_post_1ha)); qqline(qresid(f_post_1ha))
#Fitted vs. residuals
scatter.smooth(predict(f_post_1ha, type = "response"), qresid(f_post_1ha), col = "grey")
```
